{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-07-13 README.Rmd
nutrient load data for stockton and tracy
tracy discharge info to draft
c10 d7 trends plot
update clams plot with biomass flow reg

`chld7.RData` WRTDS model for chl at D7

@@

`diat_dat.RData`* additional WRTDS models for D7, including chlorphyll and sio2

@@

@@

>>>>

@@

@@

`h3dat.RData`* mean models to support third hypothesis/case study in manuscript, wrtds mean models for sio2, din, chla at c10, d7

@@

`potw_load.RData`* nutrient load data from Stockton and Tracy wwtp, kg/day

### 2016-07-13 sftrends_manu.Rnw
update clams plot with biomass flow reg

\cref{fig:clmchl}, \cref{tab:chlsio}

@@

<<clmchl, fig.height = 7, fig.width = 10, out.width = '\\textwidth', fig.cap = 'Trends in clam abundance and \\ac{chla} concentration from 1976 to 2014 at station D7 in Suisun Bay.  Invasion by \\textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \\textit{Corbicula fluminea} was shown by changes in clam density (a, annual means).  A coincident decrease in \\ac{chla} concentration was also observed (c).  A weak but significant ($p < 0.001$) relationship between clam biomass and \\ac{chla} concentration is shown in subfigure (b).  Flow relationships with \\ac{chla} concentration have also changed over time (d, observations from June). Chlorophyll shows a slight positive then dominantly negative association with increasing flow (decreasing salinity) early in the time series, whereas the trend is reversed in recent years.'>>=

# clam and model data at D7
data(clams)
data(diat_dat)
data(chld7)

# format chl data and clam data for first two plots
chldat <- filter(diat_dat, resvar %in% 'chl') %>% 
  .$mod %>% 
  .[[1]] %>% 
  annual_agg %>% 
  mutate(year = lubridate::year(date)) %>% 
  select(date, year, res, norm0.5) %>% 
  mutate(res = exp(res), norm0.5 = exp(norm0.5)) %>% 
  rename(yr = date)

@@

# subfig labs
txt <- data.frame(
  x = c(as.numeric(xlims[1]), as.numeric(xlims[2]), 25, 0.25),
  y =  c(10000, 11, 4, 3.8), 
  txt = c('a', 'c', 'b', 'd')
)
size <- 7

# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity') + 
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    # axis.title.x = element_blank(),
    legend.position = 'top', 
    legend.title = element_blank()
    ) + 
  scale_fill_brewer(
    palette = 'Set2',
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('', limits = xlims) +
  scale_y_continuous('Clams per sample') +
  with(txt[1, ], annotate('text', x = xlims[1], y = y, label = txt, size = size))

# annual chlorophyll time series with flow-norm
p2 <- ggplot(chldat, aes(x = yr, y = norm0.5)) + 
  geom_point(aes(y = res, shape = 'Observed'), size = 3) + 
  geom_line(aes(color = 'Flow-normalized')) + 
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    # axis.title.x = element_blank(), 
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) + 
  scale_colour_manual(values = c('black', 'black')) +
  scale_x_date('', limits = xlims) +
  scale_y_continuous(expression(paste("Chl-", italic(a), " (", italic(mu), "g ", L^-1, ")"))) +
  with(txt[2, ], annotate('text', x = xlims[2], y = y, label = txt, size = size))

# chlorophyll by month
chldat <- filter(diat_dat, resvar %in% 'chl') %>% 
  .$mod %>% 
  .[[1]] %>% 
  select(year, month, res, norm0.5) %>% 
  mutate(res = exp(res), norm0.5 = exp(norm0.5)) %>% 
  rename(yr = year, mo = month) %>% 
  group_by(yr, mo) %>% 
  dplyr::summarize(res = mean(res, na.rm = TRUE), norm0.5 = mean(norm0.5, na.rm = TRUE)) %>% 
  ungroup

@@

# linear model of concentration x biomass
mod <- lm(log(1 + norm0.5) ~ log(1 + biomass), data = toplo) %>% 
  coefficients
xvals <- range(toplo$biomass, na.rm = TRUE)
xvals <- seq(xvals[1], xvals[2], length = 100)
yvals <- (exp(mod[1] + mod[2] * log(1 + xvals))) - 1

xlab <- expression(paste(italic(Potamocorbula), " biomass (g ", m^-2, ")"))
ylab <- expression(paste("Flow-normalized chlorophyll (   ", italic(mu), "g ", L^-1, ")"))

# scatter plot of concetration x biomass with model 
p3 <- ggplot(toplo, aes(x = biomass, y = norm0.5)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) +
  scale_colour_manual(label = 'log(1 + chl) = 1.1 - 0.09log(1 + biomass)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0.8, 4)) +
  geom_text(data = txt[3, ], aes(x = x, y = y, label = txt, colour = NULL), show.legend = F, size = size)

# dynaplot for chl, june
cols <- RColorBrewer::brewer.pal(10, 'Set2')[1:2]
ylab <- attr(chld7, 'reslab')

p4 <- dynaplot(chld7, mo = 6, col_vec = cols, floscl = F) +
  theme_minimal() +
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal',
    strip.background = element_blank(),
    strip.text.x = element_blank()
    ) + 
  scale_x_continuous('ln-Salinity (psu)') + 
  scale_y_continuous(ylab, limits = c(0.4, 3.8)) +
  annotate('text', x = txt[4, 'x'], y = txt[4, 'y'], label = txt[4, 'txt'], size = size)

@@

# combine all
grid.arrange(
  arrangeGrob(pA, pB, ncol = 1, heights = c(1, 1)), 
  arrangeGrob(pC, pD, ncol = 1, heights = c(1, 1)), 
  ncol = 2, widths = c(1, 0.6)
  )
@

@@

pdinsy1 <- seasyrplot(toplo1$mod[[2]], col_vec = cols, logspace = F, predicted = F) +
  ptheme + 
  theme(legend.position = 'top', axis.title.y = element_blank())
pleg2 <- g_legend(pdinsy1)
pdinsy1 <- pdinsy1 + theme(legend.position = 'none')

>>>>

\cref{fig:clmchl}, \cref{fig:d7c10trnds}, \cref{tab:d7chg}, \cref{tab:c10chg}

@@

<<clmchl, fig.height = 7, fig.width = 8, out.width = '\\textwidth', fig.cap = 'Trends in clam abundance and \\ac{chla} concentration from 1976 to 2014 at station D7 in Suisun Bay.  Invasion by \\textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \\textit{Corbicula fluminea} was shown by changes in clam density (a, annual means).  A coincident decrease in \\ac{chla} concentration was also observed (c).  A weak but significant ($p < 0.001$) relationship between clam biomass and \\ac{chla} concentration is shown in subfigure (b).  Flow relationships with \\ac{chla} concentration have also changed over time (d, observations from June). Chlorophyll shows a slight positive then dominantly negative association with increasing flow (decreasing salinity) early in the time series, whereas the trend is reversed in recent years.'>>=

# clam and model data at D7
data(clams)
data(h3dat)
data(flow_dat)
chld7 <- filter(h3dat, Site_Code %in% 'D7' & resvar %in% 'chl')$mod[[1]]

@@

# subfig labs
txt <- data.frame(
  x = c(as.numeric(as.Date('1979-10-01')), 2500, 25, 0.25),
  y =  c(10000, 15, 4, 3.3), 
  txt = c('(a)', '(c)', '(b)', '(d)')
)
size <- 6

##
# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity') + 
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    # axis.title.x = element_blank(),
    legend.position = 'top', 
    legend.title = element_blank()
    ) + 
  scale_fill_brewer(
    palette = 'Set2',
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('Year', limits = xlims) +
  scale_y_continuous('Clams per sample') +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  with(txt[1, ], annotate('text', x = xlims[1] + 365 * 2, y = y, label = txt, size = size))

##
# clam biomass related to daily sacramento flow
clmdat <- group_by(clams, yr, species) %>% 
  dplyr::summarize(
    clams_smp = mean(clams_smp, na.rm = TRUE),
    biomass = mean(biomass, na.rm = TRUE)
    ) %>% 
  ungroup %>% 
  mutate(yr = as.Date(paste0(yr, '-10-01')))

flodat <- filter(flow_dat, station %in% 'sac')
clmflo <- select(clams, biomass, clams_smp, species, date) %>% 
  rename(Date = date) %>% 
  left_join(., flodat, by = 'Date') %>% 
  filter(species %in% 'Potamocorbula')

# linear model of concentration x biomass
mod <- lm(log(1 + biomass) ~ log(1 + q), data = clmflo) %>% 
  coefficients
xvals <- range(clmflo$q, na.rm = TRUE)
xvals <- seq(xvals[1], xvals[2], length = 100)
yvals <- (exp(mod[1] + mod[2] * log(1 + xvals))) - 1

ylab <- expression(paste(italic(Potamocorbula), " biomass (g ", m^-2, ")"))
xlab <- expression(paste("Sacramento daily flow (", m^3, ' ', s^-1, ")"))

# scatter plot of concentration x biomass with model 
p2 <- ggplot(clmflo, aes(x = q, y = biomass)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) +
  scale_colour_manual(label = 'log(1 + biomass) = 5.2 - 0.6log(1 + flow)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0,15)) +
  geom_text(data = txt[2, ], aes(x = x, y = y, label = txt, colour = NULL), show.legend = F, size = size)

##
# chlorophyll by month
chldat <- filter(h3dat, resvar %in% 'chl' & Site_Code %in% 'D7') %>% 
  .$mod %>% 
  .[[1]] %>% 
  select(year, month, res, bt_norm) %>% 
  mutate(res = exp(res)) %>% 
  rename(yr = year, mo = month, bt_res = res) %>% 
  group_by(yr, mo) %>% 
  dplyr::summarize(res = mean(bt_res, na.rm = TRUE), bt_norm = mean(bt_norm, na.rm = TRUE)) %>% 
  ungroup

@@

# linear model of concentration x biomass
mod <- lm(log(1 + bt_norm) ~ log(1 + biomass), data = toplo) %>% 
  coefficients
xvals <- range(toplo$biomass, na.rm = TRUE)
xvals <- seq(xvals[1], xvals[2], length = 100)
yvals <- (exp(mod[1] + mod[2] * log(1 + xvals))) - 1

xlab <- expression(paste(italic(Potamocorbula), " biomass (g ", m^-2, ")"))
ylab <- expression(paste("Flow-normalized chlorophyll (", italic(mu), "g ", L^-1, ")"))

# scatter plot of concetration x biomass with model 
p3 <- ggplot(toplo, aes(x = biomass, y = bt_norm)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) +
  scale_colour_manual(label = 'log(1 + chl) = 1.35 - 0.1log(1 + biomass)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0.8, 4)) +
  geom_text(data = txt[3, ], aes(x = x, y = y, label = txt, colour = NULL), show.legend = F, size = size)

##
# dynaplot for chl, june
cols <- RColorBrewer::brewer.pal(10, 'Set2')[1:2]
ylab <- attr(chld7, 'reslab')

p4 <- dynaplot(chld7, mo = 6, col_vec = cols, floscl = F) +
  theme_minimal() +
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal',
    strip.background = element_blank(),
    strip.text.x = element_blank()
    ) + 
  scale_x_continuous('ln-Salinity (psu)') + 
  scale_y_continuous(ylab, limits = c(0.4, 3.3)) +
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10)) + 
  annotate('text', x = txt[4, 'x'], y = txt[4, 'y'], label = txt[4, 'txt'], size = size)

@@

# combine all
grid.arrange(
  arrangeGrob(pA, pB, ncol = 1, heights = c(1, 1)), 
  arrangeGrob(pC, pD, ncol = 1, heights = c(1.05, 1)), 
  ncol = 2, widths = c(1, 1)
  )
@

@@

pdinsy1 <- seasyrplot(toplo1$mod[[2]], col_vec = cols, logspace = F, predicted = F) +
  ptheme + 
  theme(legend.position = 'top', axis.title.y = element_blank()) + 
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10))
pleg2 <- g_legend(pdinsy1)
pdinsy1 <- pdinsy1 + theme(legend.position = 'none')
